Static Discharge of Dynamic, Higher-Order Contracts

### 2013-11-14 pldi2014/content/discharge.tex
Good progress.

\subsection{Expectation that many dynamic checks are unnecessary}

Programs that can result in an error are incorrect; the purpose of blame assignment is to facilitate the act of correcting the program.
We expect a program to result in a blame assignment only rarely.
We have similar expections for a program to result in divergence, and that specifically at the hand of a chaperoned value.
In fact, we expect correct programs never to exhibit such behavior, and while chaperones serve the reality of incorrect programs, the frequency of their complaints diminishes in the long run as the programmer applies their feedback.
The nature of the programming cycle indicates that there may be some chaperones which, situated within a particular program, do not influence the behavior of the program and also that the proportion of chaperoned values in such a position may increase as the program is maintained.

Strickland et al.~\cite{strickland2012chaperones} present a chaperone erasure theorem for a calculus with vector-manipulation primitives.
This theorem establishes that a program with chaperoned values which evaluates successfully will exhibit the same behavior with all the chaperones erased--that is, with chaperone installment removed from the program.
The chaperone erasure theorem, then, makes a statement about the influence chaperones have on the meaning of programs.
We present theorems which build on this conceptual result by giving the conditions under which a chaperone installation can be omitted without changing the meaning of the program.

\subsection{Intuition behind sound discharge}

@@

A chaperone \emph{cannot} be discharged if
\begin{enumerate}
\item the negative guard causes an error to be raised,
\item the negative guard diverges,
\item the negative guard does not return the same number of results as arguments given (or one extra),
\item the extra result, if provided, is not a procedure,
\item the results of the negative guard are not, at most, chaperones of the arguments,
\item the positive guard does not accept the number of results produced by the chaperoned procedure,
\item the positive guard causes an error to be raised,
\item the positive guard diverges,
\item the positive guard does not return the same number of results as arguments given, or
\item the results of the positive guard are not, at most, chaperones of the arguments.
\end{enumerate}

For a given program execution, if neither a chaperone nor its installation raises an error or diverges, it is a candidate for \emph{discharge}, or the [omission/removal] of its installation.
It is only a candidate because its application may cause a chaperone to be installed and that chaperone may not be well-behaved.
In this case, discharging the original chaperone would prevent a meaningful chaperone from being created which would make discharge unsound.
Then, intuitively, a chaperone can be discharged if it will not raise an error or diverge  \emph{and} any chaperones it installs can be discharged.

We choose to discharge a chaperone by replacing the \scheme{chap-op} operator with \scheme{(LAMBDA (x y) x)}; discharge is accomplished by simply failing to install the chaperone at all.

This immediately raises two issues.
First, multiple chaperone installations may occur at a given application site during program execution, so we cannot, in all circumstances, selectively prevent the creation of a single chaperone.
Second, chaperone installation may not be the only operation performed by an application site.
We address both by analyzing application sites and then only those in which \scheme{chap-op} is the only operator, which we refer to as \emph{chaperone-installation sites}.
The conditions to discharge a given site are stronger than that of a chaperone installed at that site: instead of demonstrating that a single chaperone satisfies a property, we must demonstrate that every chaperone created at that site satisfies that property \emph{and} that in every chaperone installation at that site, the chaperoning procedure has the same arity [LOOK AT THIS] as the to-be-chaperoned procedure.

When put in these terms, it is clear that chaperone installment at a particular site can be discharged only if it meets these conditions for every possible execution.
These conditions seem onerous.
Indeed, in Racket, users typically access chaperone installation through a library wrapper so all chaperones in the entire program share the same installation site.
This, in turn, means that a given chaperone can be safely discharged only if \emph{all} chaperones in all executions of the program can be safely discharged.
%In other words, the program must be correct up to the encoded specification.
Fortunately, with access to the whole program, we may inline definitions which allows us to multiply the number of installment sites in the program.

\subsection{Intuition behind demotion}

@@

For instance, the program
\begin{schemedisplay}
(let ([sabotage (chap-op
                 (LAMBDA (f) f)
                 (LAMBDA (f) (chap-op f (LAMBDA (x) (raise)))))])
  (let ([g (sabotage (LAMBDA (x) x))])
    (g 1)))
\end{schemedisplay}
defines a chaperoned procedure \scheme{sabotage} which wraps higher-order arguments to fail on application, defines \scheme{g} by applying \scheme{sabotage} to the identity function, and applies \scheme{g} to \scheme{1}, which raises an error.
At no point does \scheme{g} violate chaperone-specific restrictions, so it is tempting to alter its installation site (within the guard of \scheme{sabotage}) to merely impersonate the first argument.
If done, however, the program would fail at the earlier point when \scheme{sabotage} is called which certainly differs in meaning.

@@

>>>>

Programs that can result in an error are incorrect; the purpose of blame assignment is to facilitate the act of correcting the program.
We expect a program to result in a blame assignment only rarely.
In fact, we expect correct programs never to do so, and while chaperones serve the reality of incorrect programs, the frequency of their complaints diminishes in the long run as the programmer applies their feedback.
The nature of the programming cycle indicates that there may be some chaperones which, situated within a particular program, do not influence the behavior of the program and also that the proportion of chaperoned values in such a position may increase as the program is maintained.

%Instead, the program diverges in the determination of \scheme{(even? 1)} when the first recursive call

%This program avoids the degenerate behavior of contracted recursive functions by applying the contract shallowly
%Of course, this could be avoided if the domain of \scheme{even?} was properly specified and enforced by a behavioral contract, but it is recursion which

%(letrec ([(even? odd?) (values (LAMBDA (n)
%			         (let ([(test) (= n 0)])
%			           (if test
%			               #t
%			               (let ([n-1 (- n 1)])
%			                 (odd? n-1)))))
%			       (LAMBDA (n)
%			         (let ([(test) (= n 0)])
%			           (if test
%			               #f
%			               (let ([n-1 (- n 1)])
%			                 (even? n-1))))))])
%  (let ([f (chap-op (LAMBDA (n) (+ n 1))
%                               (LAMBDA (n)
%                                 (let ([(test) (even? n)])
%                                   (if test
%                                       (values (LAMBDA (n)
%                                                 (let ([(test) (odd? n)])
%                                                   (if test
%                                                       n
%                                                       (raise))))
%                                               n)
%                                       (raise)))))])
%    (f 2)))
%by the call \scheme{(f -1)}
%where we rely on our previous definition of \scheme{letrec}.
%Even though this program results in a value, our analysis cannot discharge the chaperone since the call graph rooted at each guard procedure has a cycle in it.
%It \emph{could} be done with appeal to the size-change termination property (adapted since the standard ordering on integers is not well-founded).
%Our analysis can demote the chaperone to an impersonator, though.

Strickland et al.~\cite{strickland2012chaperones} present a chaperone erasure theorem for a calculus with chaperoned vectors.
This theorem establishes that a program with chaperoned values which evaluates successfully will exhibit the same behavior with all the chaperones erased--that is, with chaperone installment removed from the program.
In essence, the theorem bounds the influence chaperones can have on correct programs.

We are after a similar result for chaperoned procedures which will allow us to identify meaningful chaperoned values---values which influence the behavior of the program via chaperoning.
While chaperoned values do so primarily by raising an error, chaperoned procedures can affect program behavior by diverging.
While we expect that programs will seldom result in divergence, and that specifically at the hand of a chaperoned value, specifications expressed in the source language are susceptible to the same programmer errors as the program proper, which includes unintended partiality.

Consider the subtle error in the program
\begin{schemedisplay}
(letrec ([even? (LAMBDA (n)
                (if (= n 0)
                      #t
                      (odd? (- n 1))))]
         [odd? (LAMBDA (n)
                 (if (= n 1)
                     #t
                     (even? (- n 1))))])
  (let ([-> (LAMBDA (dom/c rng/c)
              (LAMBDA (f) (chaperone-procedure
                           f
                           (LAMBDA (x) (values rng/c (dom/c x))))))])
    (let ([even/c ((-> nonnegative-integer/c boolean/c) even?)]
          [odd/c  ((-> nonnegative-integer/c boolean/c) odd?)])
      (let ([f ((-> even/c odd/c) (LAMBDA (n) (+ n 1)))])
        (f 1))))
\end{schemedisplay}
manifested by the body call \scheme{(f 1)}.
The programmer expects the domain violation of \scheme{f} to be reported, and it would be but for the base case logical error in \scheme{odd?}.
Instead, the numeric value passes a guard for the first and only time through \emph{even/c}, the contracts on \scheme{even?} and \scheme{odd?} being only shallowly installed.
(Shallow installation is typical of intra-module procedures---especially mutually-recursive ones.)
As a result, the logical error in \scheme{odd?} is not caught by the contract system, despite it being the precise situation for which the system was designed, and the program diverges unchecked.\footnote{The programmer could embed the specifications deeper into the definitions but would then pay the well-known contract penalty.
As long as contracts remain expensive, cheaper alternatives but nevertheless with high cost--such as the ``remote'' possibility of incorrectness--prevail.}

This example illustrates the need to include divergence in the set of effects a judgment of meaningfulness must consider.

\subsection{Sound chaperone discharge}

If a specific chaperoned value is not meaningful in a particular execution, it can be soundly \emph{discharged}, meaning that it is safe to replace its particular installation with identity if it is possible to do so.
(The soundness claim does not say it is necessarily possible---only safe---to do so.)
A chaperone is meaningful if
\begin{enumerate}
\item the negative guard causes an error to be raised,
\item \label{con:neg-divergence} the negative guard diverges,
\item the negative guard does not return the same number of results as arguments given (or one extra),
\item the extra result, if provided, is not a procedure,
\item the results of the negative guard are not, at most, chaperones of the arguments,
\item the positive guard does not accept the number of results produced by the chaperoned procedure,
\item the positive guard causes an error to be raised,
\item the positive guard diverges,
\item the positive guard does not return the same number of results as arguments given,
\item the results of the positive guard are not, at most, chaperones of the arguments, or
\item its application effects the installation of a meaningful chaperone.
\end{enumerate}

We discharge a chaperone by replacing the operator at its point of installation with \scheme{(LAMBDA (x y) x)}; discharge is accomplished by discarding the guard procedure in lieu of attaching it.

This strategy immediately shifts our focus from dynamic points of installation to static points of installation in the following way:
First, chaperone installation may not be the only operation performed at a call site.
Thus, we must restrict ourselves to call sites dedicated to that purpose.
Second, even once we have restricted ourselves to bonafide chaperone-installation sites, there may be more than one chaperone installed at each site.
Thus, we must restrict ourselves to sites at which no meaningful chaperone is installed.
Finally, we must extend these conditions from a particular execution to all possible executions.
% multiple operations performed at a particular site
% multiple chaperones installed at a particular site (during one particular execution)
% the chaper
First, multiple chaperone installations may occur at a given application site during program execution, so we cannot, in all circumstances, selectively prevent the creation of a single chaperone.

to points of installation

This strategy immediately raises three issues.
Such a strategy affects all executions

Using this approach to discharge, it is clear that chaperone installment at a particular site can be discharged only if it meets these conditions for every possible execution.
These conditions seem onerous.
Indeed, in Racket, users typically access chaperone installation through a library wrapper so all chaperones in the entire program share the same installation site.
This, in turn, means that a given chaperone can be safely discharged only if \emph{all} chaperones in all executions of the program can be safely discharged.
%In other words, the program must be correct up to the encoded specification.
Fortunately, with access to the whole program, we may inline definitions which allows us to multiply the number of installment sites in the program.

This immediately raises two issues.

Second, chaperone installation may not be the only operation performed by an application site.
We address both by analyzing application sites and then only those in which \scheme{chap-op} is the only operator, which we refer to as \emph{chaperone-installation sites}.
The conditions to discharge a given site are stronger than that of a chaperone installed at that site: instead of demonstrating that a single chaperone is not meaningful, we must demonstrate that every chaperone created at that site is not meaningful \emph{and} that the chaperone installations themselves are not meaningful.

% in every chaperone installation at that site, the chaperoning procedure has the same arity [LOOK AT THIS] as the to-be-chaperoned procedure.

@@

For a given program execution, if neither a chaperone nor its installation raises an error or diverges, it is a candidate for \emph{discharge}, or the targeted prevention of its installation.
It is only a candidate because its application may cause a meaningful chaperone to be installed, i.e., a chaperone that is not so well-behaved.
In this case, discharging the original chaperone would prevent a meaningful chaperone from being created which would make discharge unsound.
Then, intuitively, a chaperone can be discharged if it will not raise an error or diverge  \emph{and} any chaperones it installs can be discharged.

\subsection{Sound chaperone demotion}

@@

For instance, the program
\begin{schemedisplay}
(let ([sabotage (chap-op (LAMBDA (f) f) (LAMBDA (f) (chap-op f raise)))])
  (let ([g (sabotage (LAMBDA (x) x))])
    (g 1)))
\end{schemedisplay}
defines a chaperoned procedure \scheme{sabotage} which wraps higher-order arguments to fail on application, defines \scheme{g} by applying \scheme{sabotage} to the identity function, and applies \scheme{g} to \scheme{1}, which raises an error.
At no point does \scheme{g} violate chaperone-specific restrictions, so it is tempting to alter its installation site (within the guard of \scheme{sabotage}) to merely impersonate the first argument.
If done, however, the program would fail at the earlier point when \scheme{sabotage} is called which certainly differs in meaning.

@@

\subsection{Dependence}

Harrison~\cite{harrison1989interprocedural} characterizes the principle of dependence of procedures on resources.
It states, roughly, that, in the absence of tail-call optimization, any procedures with a live frame on the stack at a point of resource access \emph{depend} on that resource.
%A definition of dependence without the tail-call optimization assumption is incomplete as there could be procedures which haven't returned (and should still be considered active from the perspective of dependence) but will not return as they are subsumed by a tail call.
Assuming degenerate tail-call behavior simply ensures that every procedure will return explicitly so that the stack represents a complete set of dependent procedures.

We reverse the notion of dependence to say that certain control events depend on the active procedures on the stack.
With our ultimate goal being to prevent chaperone installation, we are only concerned with the chaperone guard procedures depended upon at each point of interest.
Even in the presence of tail-call optimization, as in \chapcalc, the presence of active chaperone guard procedures is always recorded on the stack, and we can apply Harrison's dependence reasoning fairly directly.

%Dependence for our purposes here is defined in terms of active chaperoned operators.
%Specifically, we are interested in which chaperone guard procedures are currently active and we reverse the notion of dependence to

%The assumption of degenerate tail-call behavior ensures that all procedures which haven't returned are included
%We base our notion of dependence on the principle of dependence of Harrison~\cite{harrison1989interprocedural} which, roughly, is that procedures active at the point of resource access are \emph{dependent} on that resource.
%This principle assumes the absence of tail-call optimization; a procedure depends on the resource even if its activation frame was subsumed by a tail call.
%We reverse the relationship before applying it to chaperone installation and error assertion: the creation of a chaperone or assertion of an error depends on all procedures active at that point.
%Our calculus ``optimizes'' tail calls and we consider procedures subsumed to be dependent also.
%However, we are concerned only with which chaperone guards are dependent on one of these things, and they leave evidence of their invocation on the stack--even in the presence of proper tail-call optimization--until their continuation is accessed.

### 2013-11-14 pldi2014/content/interposition.tex
Progress.
Good progress.

There are three distinct phases in the contract lifecycle:
The \emph{creation} phase occurs when the application of contract combinators is evaluated---for instance \scheme{(number? . -> . number?)}.
The \emph{projection} phase occurs when the contract is attached to a value and so may occur multiple times.
We can imagine a procedure which maps a simple contract and higher-order value to a contracted higher-order value which would behave similar to
\begin{schemedisplay}
(define (-> dom/c rng/c)
  (LAMBDA (f)
    (LAMBDA (x)
      (rng/c (f (dom/c x))))))
\end{schemedisplay}
The \emph{application} phase occurs whenever the contract is checked.
For contracted procedures, the domain and range contracts are checked whenever the procedure is applied.

@@

The contracted tail-recursive factorial procedure expressed as a chaperone is
\begin{schemedisplay}
(define fact (chaperone-procedure
              (LAMBDA (n a)
                (if (= n 0)
                    a
                    (fact (- n 1) (* n a))))
              (LAMBDA (n a)
                (if (and (nonnegative-integer? a)
                         (positive-integer? a))
                    (values (LAMBDA (a)
                              (if (positive-integer? a)
                                  a
                                  (raise positive blame ...)))
                            n a)
                    (raise negative blame ...)))))
\end{schemedisplay}

@@

Thus, we would prefer to remove all application costs due to interposition, but will accept merely removing the costs specific to chaperoning.

>>>>

There are three distinct phases in the contract lifecycle:
The \emph{creation} phase occurs when the application of contract combinators is evaluated---for instance \scheme{(number? . -> . number?)}.
The \emph{projection} phase occurs when the contract is attached to a value and so may occur multiple times.
We can imagine a procedure which maps a simple contract and higher-order value to a contracted higher-order value which would behave similar to
\begin{schemedisplay}
(define (-> dom/c rng/c)
  (LAMBDA (f) (LAMBDA (x) (rng/c (f (dom/c x))))))
\end{schemedisplay}
The \emph{application} phase occurs whenever the contract is checked.
For contracted procedures, the domain contract is checked whenever the procedure is applied and the range contract when the result is obtained.

@@

From the view of impersonators, contract ascription to the tail-recursive factorial can be expressed by
\begin{schemedisplay}
(define fact (chaperone-procedure
              (LAMBDA (n a)
                (if (= n 0)
                    a
                    (fact (- n 1) (* n a))))
              (LAMBDA (n a)
                (if (and (nonnegative-integer? a)
                         (positive-integer? a))
                    (values (LAMBDA (a)
                              (if (positive-integer? a)
                                  a
                                  (raise positive blame ...)))
                            n a)
                    (raise negative blame ...)))))
\end{schemedisplay}

@@

Thus, we would prefer to remove all application costs due to interposition, but, when unattainable, will accept merely removing the costs specific to chaperoning.

### 2013-11-14 pldi2014/content/introduction.tex
Good progress.

Consider, for instance, the program
\begin{schemedisplay}
(define/contract (fact n a)
  (nonnegative-integer/c positive-integer/c . -> . positive-integer/c)
  (if (= n 0)
      a
      (fact (- n 1) (* n a))))
\end{schemedisplay}
which attaches dynamic domain and range contracts to the tail-recursive factorial.
This contracted procedure behaves as if it were written
\begin{schemedisplay}
(define (fact n a)
  (positive-integer/c
   ((LAMBDA (n a)
      (if (= n 0)
          a
          (fact (- n 1) (* n a))))
    (nonnegative-integer/c n)
    (positive-integer/c a))))
\end{schemedisplay}
where \scheme{positive-integer/c} and \scheme{nonnegative-integer/c} are procedures which raise an error if their argument does not belong to the domain reflected by the procedure's name.
Written in this way, it is apparent that \scheme{fac} is not tail-recursive.
Instead, the context grows with every recursive call so that \scheme{(fact 6 1)} reduces as
\begin{schemedisplay}
(fact 6 1)
(positive-integer/c (fact 5 6))
(positive-integer/c (positive-integer/c (fact 4 30)))
...
(positive-integer/c ... (positive-integer/c (fact 0 720)))
(positive-integer/c ... (positive-integer/c 720))
\end{schemedisplay}
whence the result will be validated six times---and \scheme{n} times for \scheme{(fact n 1)}.
The attachment of these contracts has transformed a constant-space algorithm into a linear-space one and, while the time complexity hasn't suffered asymptotically, the accumulated checks represent a real (and, in this case, unnecessary) cost.
Thus, a simple contract ascription can upset our performance expectations for even simple procedures.

As contract specifications become more elaborate, they become more susceptible to error.
In contract systems, such as Racket's~\cite{plt-tr1}, which allow arbitrary predicate procedures written in the source language itself to serve as specifications, this error is effected by the programmer.
It is only natural, then, to contract these procedures as well, an act which aides error diagnosis but compounds the run-time cost of contracts.
In these systems, execution time of the program proper can be dwarfed by that of the enforcement of its specification (in one case accounting for 96\% of total execution time) \cite{strickland2012chaperones}.

@@

Recent advances in higher-order program analysis \cite{earl2012introspective} provide perfectly precise return flow and allow the form of the stack to be approximated at each control point in the program.
The original purpose of this approximation was to provide an abstract garbage collector a conservative set of live frames.
We repurpose this capability to analyze a notion of dependence similar to that of Might and Prabhu~\cite{might2009interprocedural} allowing us to judge the safe removal of dynamic contracts \emph{even when the installation of other contracts depends on them}.
The expressiveness of Racket's contract language admits dynamic checks which diverge, so we must integrate a termination analysis based on Sereni's~\cite{sereni2007termination} that informs the discharge of dependently-installed contracts.

We proceed as follows:
We discuss interposition and its relationship to behavioral contracts in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for interposing higher-order values, in Section 3.
We present two theorems which establish the conditions for the safe removal or weakening of interposition in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We discuss the influence of parameters of the analysis on precision and performance in Section 6.
We discuss some Racket language features which require more care to integrate with the present analysis in Section 7.
We conclude by discussing the future application and direction of this work in Section 8.

>>>>

Consider, for instance, the program
\begin{schemedisplay}
(define/contract (fact n a)
  (nonnegative-integer/c positive-integer/c . -> . positive-integer/c)
  (if (= n 0)
      a
      (fact (- n 1) (* n a))))
\end{schemedisplay}
which attaches dynamic domain and range contracts to the tail-recursive factorial.
This contracted procedure behaves as if it were written
\begin{schemedisplay}
(define (fact n a)
  (positive-integer/c
   ((LAMBDA (n a)
      (if (= n 0)
          a
          (fact (- n 1) (* n a))))
    (nonnegative-integer/c n)
    (positive-integer/c a))))
\end{schemedisplay}
where \scheme{positive-integer/c} and \scheme{nonnegative-integer/c} are procedures which raise an error if their argument does not belong to the domain reflected by the procedure's name.
Written in this way, it is apparent that \scheme{fact} is not tail-recursive.
Instead, its context grows with every recursive call so that \scheme{(fact 6 1)} reduces as
\begin{schemedisplay}
(fact 6 1)
(positive-integer/c (fact 5 6))
(positive-integer/c (positive-integer/c (fact 4 30)))
...
(positive-integer/c ... (positive-integer/c (fact 0 720)))
(positive-integer/c ... (positive-integer/c 720))
\end{schemedisplay}
whence the result will be validated six times---and \scheme{n} times for \scheme{(fact n 1)}.
The attachment of these contracts has transformed a constant-space algorithm into a linear-space one and, while the time complexity hasn't suffered asymptotically, the accumulated checks represent a real (and, in this case, unnecessary) cost.
Thus, a simple contract ascription can upset our performance expectations for even simple procedures.

As contract specifications become more elaborate, they become more susceptible to error.
In contract systems, such as Racket's~\cite{findler2002contracts}, which allow arbitrary predicate procedures written in the source language itself to serve as specifications, this error is effected by the programmer.
It is only natural, then, to contract these procedures as well, an act which aides error diagnosis but compounds the run-time cost of contracts.
In these systems, execution time of the program proper can be dwarfed by that of the enforcement of its specification (in one case accounting for 96\% of total execution time) \cite{strickland2012chaperones}.

@@

Recent advances in higher-order program analysis \cite{earl2012introspective} provide perfectly precise return flow and allow the form of the stack to be approximated at each control point in the program.
The original purpose of this approximation was to provide an abstract garbage collector a conservative set of live frames.
We repurpose this capability to analyze a notion of dependence similar to that of Harrison~\cite{harrison1989interprocedural} allowing us to judge the safe prevention of dynamic contract installation---that is, contracts \emph{installed} at run time---\emph{even when the installation of other contracts depends on them}.
The expressiveness of Racket's contract language admits dynamic checks which diverge, so we must integrate a termination analysis based on Sereni's~\cite{sereni2007termination} that informs the discharge of dependently-installed contracts.

We proceed as follows:
We discuss interposition and its relationship to behavioral contracts in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for interposing higher-order values, in Section 3.
We present two theorems which establish the conditions for the safe removal or weakening of interposition in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We discuss the influence of parameters of the analysis on precision and performance in Section 6.
%We discuss some Racket language features which require more care to integrate with the present analysis in Section 7.
We discuss related work in Section 7.
We conclude by discussing the integration of some Racket language features with this analysis in Section 8.
%future application and direction of this work in Section 8.

### 2013-11-14 pldi2014/content/limitations.tex
Progress.
Good progress.

\section{Limitations}

Our calculus has an impoverished value domain, so our termination judgments of guards are nearly complete in practice as a result of only a simple flow analysis.
(Because, in practice, guards are not recursive themselves or any procedures in the call graph. Essentially, call graphs for guard procedures are trees.)
Contracts on compound data complicate matters because the data may be defined inductively which makes recursive checking natural.
We can accomodate this in particular by integrating size-change analysis~\cite{lee2001size} into our abstract interpretation.
This would be novel as higher-order size-change analysis~\cite{sereni2007termination} has heretofore been applied only to variants of $k$-CFA~\cite{shivers1991control}.

\scheme{app-values} increases the expressiveness of the language.
(We do not have to specify the number of arguments to a function statically.)
The structure of a list is recursive.
We can untie the recursive definition by directing through the store; we took this approach with environments.
We could look into this, but that means the length of the list could vary.
That could \emph{really} destroy precision, but we need to investigate it further.

\emph{rest values} increase the expressiveness of the language. They are very related to \scheme{app-values}.
\scheme{chap-op} as a primitive (as opposed to a form) increases the expressiveness of the language.

Our analysis applies equally well if we extend the language with more referentially transparent primitives or macro-expressible constructs such as \scheme{and} and \scheme{or} forms.
We can even add an error-catching facility but we must be careful in our analysis.

@@

Chaperones are intended to be treated like the original value.
% box?, procedure?, etc. will respond the same way
% they evaluate equal? to the values they wrap, but not eq?
Procedure chaperones are considered to have the same arity of the procedures they wrap.
The \scheme{chaperone?} predicate is designated as the single way to detect chaperones at runtime, though its use is discouraged.

\subsection{Handling \emph{handle}}

Exception handlers prevent certain errors from escaping a particular dynamic extent.
Consider the program
\begin{schemedisplay}
(handle
  (chap-op + (raise) (raise))
  -)
\end{schemedisplay}
wherein an error is always raised and the handler always triggered.
Even though an exception never escapes the dynamic extent of the \scheme{handle} expression, replacing the \scheme{chap-op} application with its first argument results in a different program. 
This means that we must consider whether a \scheme{chap-op} application ever raises an error, and not simply if that error escapes a delimited extent.
% being able to catch errors means that errors can change the meaning of a program arbitrarily--even if it results in a value.
% this means that we need to preserve errors
% chaperone-of? within the language can change the behavior, right? if it was a chaperone before, removing the chaperone makes it still chaperone-of? it's chaperone? that makes it possible to detect changes, but that is discouraged, and we will leave it out of our language.

chaperone and impersonator properties

>>>>

\section{Limitations}
\label{sec:limitations}

There is a litany of Racket features with which a complete analysis must contend.

Our calculus has an impoverished value domain, so our termination judgments of guards are nearly complete in practice as a result of only a simple flow analysis.
%Because, in practice, guards are not recursive themselves or any procedures in the call graph. Essentially, call graphs for guard procedures are trees.)
Contracts on compound data complicate matters because the data may be defined inductively which makes recursive checking natural.
We can accomodate this in particular by integrating size-change analysis~\cite{lee2001size} into our abstract interpretation.
This would be novel as higher-order size-change analysis~\cite{sereni2007termination} has heretofore been applied only to variants of $k$-CFA~\cite{shivers1991control}.

Racket allows first-class \emph{impersonator properties} to be defined with which the programmer can associate arbitrary values with impersonator values and query whether a given value is an impersonator with such a property.
We have completely ignored this aspect of impersonators in our analysis.
Such properties and their associated values must be included at the installation site for the impersonator, so we can safely ignore them by only considering sites at which no properties are installed.

Of course, Racket provides impersonation facilities for \emph{many} other classes of values: boxes, structures, vectors, hashes, and even continuation prompts.
Procedures are ultimately one of the simpler datatypes having only one operation common among them.
We reserve investigation into the analysis of other datatypes for future work.

Other features inherited from Scheme are mutation, variable-arity procedures.

On its face, the integration of an error-handling construct is tractable, at least with the current, limited calculus.
We simply need to keep in mind that errors influence the behavior of the program whether they are redirected or not.
Essentially, we would treat exception handlers for error-dependence analysis as if they didn't exist.
%mutation
%different types of impersonators
%impersonator properties

%\scheme{app-values} increases the expressiveness of the language.
%(We do not have to specify the number of arguments to a function statically.)
%The structure of a list is recursive.
%We can untie the recursive definition by directing through the store; we took this approach with environments.
%We could look into this, but that means the length of the list could vary.
%That could \emph{really} destroy precision, but we need to investigate it further.

%\emph{rest values} increase the expressiveness of the language. They are very related to \scheme{app-values}.
%\scheme{chap-op} as a primitive (as opposed to a form) increases the expressiveness of the language.

%Our analysis applies equally well if we extend the language with more referentially transparent primitives or macro-expressible constructs such as \scheme{and} and \scheme{or} forms.
%We can even add an error-catching facility but we must be careful in our analysis.

@@

Chaperones are intended to be treated like the original value.
The \scheme{chaperone?} predicate is designated as the single way to detect chaperones at runtime, though its use is discouraged.

%\subsection{Handling \emph{handle}}

%Exception handlers prevent certain errors from escaping a particular dynamic extent.
%Consider the program
%\begin{schemedisplay}
%(handle
%  (chap-op + (raise) (raise))
%  -)
%\end{schemedisplay}
%wherein an error is always raised and the handler always triggered.
%Even though an exception never escapes the dynamic extent of the \scheme{handle} expression, replacing the \scheme{chap-op} application with its first argument results in a different program. 
%This means that we must consider whether a \scheme{chap-op} application ever raises an error, and not simply if that error escapes a delimited extent.
% being able to catch errors means that errors can change the meaning of a program arbitrarily--even if it results in a value.
% this means that we need to preserve errors
% chaperone-of? within the language can change the behavior, right? if it was a chaperone before, removing the chaperone makes it still chaperone-of? it's chaperone? that makes it possible to detect changes, but that is discouraged, and we will leave it out of our language.

%chaperone and impersonator properties
